### 2010-02-18 doc/tech_doc/a-rex/arex_tech_doc.tex
Switching documents to bibliography style which is supposed to be present in latex distributions. Fixes in some documents and build procedures.
add info on running a-rex under GridFTP interface

\documentclass{article}                            %for shorter notes
\usepackage{graphics}                              %for EPS images (latex)
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{listings}                              %for code listings
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\lstloadlanguages{bash,XML}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {A.Konstantinov},
  pdftitle = {The ARC Computational Job Management Component -- A-REX},
  pdfsubject = {},
  pdfkeywords = {},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
\definecolor{lightblue}{cmyk}{0.05,0,0,0.05}
\lstset{
frame=single,
backgroundcolor=\color{shadecolor},
fillcolor=\color{shadecolor},
rulecolor=\color{lightblue},
basicstyle=\ttfamily,
xleftmargin=0.05\linewidth,
aboveskip=2\medskipamount
}
%
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

%-------- Change this to NORDUGRID-XXXXXXX-NN

@@

%\tableofcontents                          %Comment if use article style
\newpage
\section{Introduction\label{sec:intro}}

A-REX is one of ARC middleware components that implements functions
of the so-called \emph{Computing Element} (CE). Here Computing Element
is a service accepting requests containing description of generic
computational job and executing it in the underlying local batch system.

It takes care of job pre- and post-processing. This comprises
stage-in of files containing input data or program modules from wide
range of sources and transfer or storing of output results.

A-REX implements a Web Service (WS) interface which provides a
way to submit and control computational tasks (jobs) to be executed
by A-REX and the underlying Local Resource Management System.

\begin{framed}You should use this document for advanced configuration purposes
and understanding of internals of the aforementioned tools. For general
installation and configuration of whole system refer to other documents
available at \url{http://www.nordugrid.org/papers.html}.
\end{framed}

@@

Each job gets a directory on the CE called the \emph{session
  directory} (SD). Input files are gathered in the SD. The job may
also produce new data files in the SD. The A-REX does not guarantee
the availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environment).
The SD is also the only place which is controlled by the A-REX. It is
accessible by the user from outside through the HTTP(S) protocol.  Any
file created outside the SD is not controlled by the A-REX. Any
exchange of data between client and A-REX (including also program
modules) is performed via HTTP(S). An URL for accessing input/output
files is obtained through the WS Local Information Description
Interface (LIDI) of A-REX.

@@

At the start of a download, a dummy file with the same name and size
as the source file is created in the SD to check that enough space is
available. As the transfer proceeds this dummy file is over-written
with the real data. If the file is to be cached (see
Section~\ref{sec:cache}) then the pre-allocation of space is not done,
as cache space is managed by the A-REX. In the case of caching two
extra validation checks are also performed: firstly, if the source is
an Indexing Service then metadata (file size, checksum) reported by
the Indexing Service and the service hosting the physical replica are
compared, if available. If they differ then that replica is not
downloaded. Secondly, on completion of the download the file size is
compared to that reported by the source - if they differ then the
download is failed.

@@

From the point of view of the A-REX a job passes through various states.
Figure~\ref{job states diagram} presents a diagram of the possible
states of a job.%

@@

\item \textbf{Executing} -- The job is queued or being executed in the
  LRMS.  The A-REX takes no actions except waiting until the job
  finishes.

\item \textbf{Killing} -- Necessary action to cancel the job in the
  LRMS is being taken.

@@

In case of a failure special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by the
  user}. No files will be moved to their destination SE.

@@

In full installation the A-REX and it's components support the following
data transfer protocols and corresponding URLs: \emph{ftp, gsiftp,
http, https, lfc, rls} and \emph{srm.} For more information please see
{}``The Hosting Environment of the Advanced Resource Connector middleware''
document~\cite{hed}.

@@

For each local UNIX user listed in the A-REX configuration -- including
a generic one which covers all local user identities -- a \textit{control
directory} exists. In this directory the A-REX stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In most common configuration case A-REX
serves all users defined by the Operating System and stores their control
files in the same directory. To make it easier to recover in case of failure,
the A-REX stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\item \textit{job.ID.failed} -- the existence of this file marks the failure
of the job. It can also contain one or more lines of text describing
the reason of failure. Failure includes the return code different
from zero of the job itself.
\item \textit{job.ID.errors} -- this file contains the output produced by
external utilities like \textbf{\textit{downloader}}, \textbf{\textit{uploader}},
script for job submission to LRMS, etc on their stderr handle. Those
are not necessarily errors, but can be just useful information about
actions taken during the job processing. In case of problem include
content of that file while asking for help.
\item \textit{job.ID.diag} -- information about resources used during execution
of job and other information suitable for diagnostics and statistics.
It's format is similar to that of \textit{job.ID.local}. The following
names are at least defined:
\begin{itemize}
\item \textit{nodename} -- name of computing node which was used to execute
job,
\item \textit{runtimeenvironments} -- used runtime environments separated
by ';',
\item \textit{exitcode} -- numerical exit code of job,
\item \textit{frontend\_distribution} -- name and version of operating system
distribution on frontend computer,
\item \textit{frontend\_system} -- name of operating on frontend computer,
\item \textit{frontend\_subject} -- subject (DN) of certificate representing
frontend computer,
\item \textit{frontend\_ca} -- subject (DN) of issuer of certificate representing
frontend computer,
\end{itemize}
and other information provided by GNU \emph{time} utility. Note that
some implementation of \emph{time} insert unrequested information
in their output. Hence some lines can have broken format.
\item \textit{job.ID.proxy} -delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} -- temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different form job owner's \emph{id}.
\end{itemize}

@@

The \emph{MigrateActivity} operation generates a request to migrate a 
grid job from another ARC1 cluster, i.e. the operation will get input 
files and possibly job description from the cluster currently holding 
the job and create the job as a new activity at the present cluster.
Currently only migration of queuing jobs is supported.

@@

\begin{itemize}
\item \textit{control directory} -- In this directory A-REX stores an information
about accepted jobs. It must have full permissions there.\\
A subdirectory called \textit{logs} is created there. It is used to
accumulate information about started and finished jobs. This information
is periodically sent to the desired \emph{logger service}(s). For
each job start and stop event, and for each logger service where that
event must be sent, a separate file is written. Once an event is sent,
the corresponding file is deleted.\\
Some utility files, which contain for example cached information on remote
services, may also be found in the control directory. These files have
a \textit{.conf} suffix.
\end{itemize}

@@

Due to historical reasons configuration of the A-REX is split into 2
parts. The small part is located inside HED configuration (see
Appendix~\ref{annex:arex-conf} for schema and description of supported
elements). This configuration refers to the file containing the main
part of the configuration. The default location of the main
configuration file is \textit{/etc/arc.conf}.

@@

The A-REX is installed as an ARC1 component of ARC middleware and is available
in NorduGrid download area at ARC1 Download \url{http://download.nordugrid.org/software/nordugrid-arc1/}.There are packages available for various Linux distributions. The
A-REX comes in the \texttt{nordugrid-arc1-arex} package. Source code ready for
compilation is available too.

@@

\subsection{Setup of the A-REX}

@@

\subsection{Usage}

@@

>>>>

\documentclass{article}                            %for shorter notes
\usepackage{graphics}                              %for EPS images (latex)
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
\usepackage{listings}                              %for code listings
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\lstloadlanguages{bash,XML}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{array}
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {A.Konstantinov},
  pdftitle = {The ARC Computational Job Management Component -- A-REX},
  pdfsubject = {},
  pdfkeywords = {},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
\definecolor{lightblue}{cmyk}{0.05,0,0,0.05}
\lstset{
frame=single,
backgroundcolor=\color{shadecolor},
fillcolor=\color{shadecolor},
rulecolor=\color{lightblue},
basicstyle=\ttfamily,
xleftmargin=0.05\linewidth,
aboveskip=2\medskipamount
}
%
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

@@

\tableofcontents                          %Comment if use article style
\newpage
\section{Introduction\label{sec:intro}}

The A-REX is an ARC middleware component that implements functions
of the so-called \emph{Computing Element} (CE). Here Computing Element
is a service accepting requests containing a description of generic
computational jobs and executing it in the underlying local batch system.

It takes care of job pre- and post-processing, i.e.\ stage-in of
files containing input data or program modules from a wide range of
sources and transfer or storing of output results.

In previous versions of ARC, computational tasks (jobs) were submitted
to a resource through a GridFTP service running on the CE, and
processed by a Grid Manager (GM). The A-REX replaces both these
components, by implementing a Web Service (WS) interface which
provides a way to submit jobs, and a revised version of the GM to
process jobs and interact with the underlying Local Resource
Management System. The A-REX can also be set up to process jobs
submitted through the traditional ARC GridFTP service, as well as or
instead of those using the WS interface. Unless stated otherwise, the
remainder of this document assumes running only the WS interface.

\begin{framed}You should use this document for advanced configuration purposes
and understanding of the internals of the aforementioned tools. For
general installation and configuration of the whole system please
refer to other documents available at
\url{http://www.nordugrid.org/papers.html}.
\end{framed}

@@

Each job gets a directory on the CE called the \emph{session
  directory} (SD). Input files are gathered in the SD. The job may
also produce new data files in the SD. The A-REX does not guarantee
the availability of any other places accessible by the job other than
SD (unless such a place is part of a requested Runtime Environment).
The SD is also the only place which is controlled by the A-REX. It is
accessible by the user from outside through the HTTP(S) protocol.  Any
file created outside the SD is not controlled by the A-REX. Any
exchange of data between client and A-REX (including also program
modules) is performed via HTTP(S). A URL for accessing input/output
files is obtained through the WS Local Information Description
Interface (LIDI) of A-REX.

@@

At the start of a download initiated by the A-REX, a dummy file with
the same name and size as the source file is created in the SD to
check that enough space is available. As the transfer proceeds this
dummy file is over-written with the real data. If the file is to be
cached (see Section~\ref{sec:cache}) then the pre-allocation of space
is not done, as cache space is managed by the A-REX. In the case of
caching two extra validation checks are also performed: firstly, if
the source is an Indexing Service then metadata (file size, checksum)
reported by the Indexing Service and the service hosting the physical
replica are compared, if available. If they differ then that replica
is not downloaded. Secondly, on completion of the download the file
size is compared to that reported by the source - if they differ then
the download is failed.

@@

From the point of view of the A-REX a job passes through various states.
Figure~\ref{job states diagram} presents a diagram of the possible
states of a job.

@@

\item \textbf{Executing (InLRMS)} -- The job is queued or being executed in the
  LRMS.  The A-REX takes no actions except waiting until the job
  finishes.

\item \textbf{Killing (Canceling)} -- Necessary action to cancel the job in the
  LRMS is being taken.

@@

In case of failure, special processing is applied to output files.
All specified output files are treated as \textbf{downloadable by the
  user}. No files will be moved to their destination SE.

@@

In a full installation, the A-REX and its components support the
following data transfer protocols and corresponding URLs: \emph{ftp,
  gsiftp, http, https, lfc, rls} and \emph{srm.} For more information
please see ``The Hosting Environment of the Advanced Resource
Connector middleware'' document~\cite{hed}.

@@

For each local UNIX user listed in the A-REX configuration -- including
a generic one which covers all local user identities -- a \textit{control
directory} exists. In this directory the A-REX stores information
about jobs belonging to that user. Multiple users can share the same
\textit{control directory}. In the most common configuration case, the A-REX
serves all users defined by the Operating System and stores their control
files in the same directory. To make it easier to recover in case of failure,
the A-REX stores most information in files rather than in memory.
All files belonging to the same job have names starting with \textbf{job.ID.,}
where ID is the job identifier.

@@

\item \textit{job.ID.failed} -- the existence of this file marks the failure
of the job. It can also contain one or more lines of text describing
the reason of failure. Failure includes the return code different
from zero of the job itself.
\item \textit{job.ID.errors} -- this file contains the output produced by
external utilities like \textbf{\textit{downloader}}, \textbf{\textit{uploader}},
script for job submission to LRMS, etc on their stderr handle. Those
are not necessarily errors, but can be just useful information about
actions taken during the job processing. In case of problem include
content of that file while asking for help.
\item \textit{job.ID.diag} -- information about resources used during execution
of job and other information suitable for diagnostics and statistics.
It's format is similar to that of \textit{job.ID.local}. The following
names are at least defined:
\begin{itemize}
\item \textit{nodename} -- name of computing node which was used to execute
job,
\item \textit{runtimeenvironments} -- used runtime environments separated
by ';',
\item \textit{exitcode} -- numerical exit code of job,
\item \textit{frontend\_distribution} -- name and version of operating system
distribution on frontend computer,
\item \textit{frontend\_system} -- name of operating on frontend computer,
\item \textit{frontend\_subject} -- subject (DN) of certificate representing
frontend computer,
\item \textit{frontend\_ca} -- subject (DN) of issuer of certificate representing
frontend computer,
\end{itemize}
and other information provided by GNU \emph{time} utility. Note that
some implementations of \emph{time} insert unrequested information
in their output. Hence some lines can have broken format.
\item \textit{job.ID.proxy} -- delegated X509 credentials.
\item \textit{job.ID.proxy.tmp} -- temporary X509 credentials with different
UNIX ownership used by processes run with effective \emph{user id}
different from job owner's \emph{id}.
\end{itemize}

@@

The \emph{MigrateActivity} operation generates a request to migrate a 
grid job from another A-REX, i.e. the operation will get input 
files and possibly job description from the cluster currently holding 
the job and create the job as a new activity at the present cluster.
Currently only migration of queuing jobs is supported.

@@

\begin{itemize}
\item \textit{control directory} -- In this directory A-REX stores information
about accepted jobs. It must have full permissions there.\\
A subdirectory called \textit{logs} is created there. It is used to
accumulate information about started and finished jobs. This information
is periodically sent to the desired \emph{logger service}(s). For
each job start and stop event, and for each logger service where that
event must be sent, a separate file is written. Once an event is sent,
the corresponding file is deleted.\\
Some utility files, which contain for example cached information on remote
services, may also be found in the control directory. These files have
a \textit{.conf} suffix.
\end{itemize}

@@

Due to historical reasons, configuration of the A-REX is split into 2
parts: HED configuration for the WS interface and a legacy
configuration file for the rest. For more information on the HED
configuration see instructions in Section~\ref{sec:wssetup}, and
Appendix~\ref{annex:arex-conf} for a full schema and description of
supported elements. The HED configuration refers to the legacy file,
which is identical to the configuration file used for the GridFTP/Grid
Manager services in previous versions of ARC. The default location of
this legacy file is \textit{/etc/arc.conf}. If the A-REX is run
without the WS interface, then only the legacy file is needed (to
configure the GridFTP interface see~\cite{gm}). This section describes
the content of the legacy file relating to running the A-REX.

@@

The A-REX is installed as a component of the ARC middleware and
packages for various distributions are available from NorduGrid
repositories, or directly from the download area of the NorduGrid
website. Source code ready for compilation is available too.

@@

\subsection{Setup of the A-REX with WS Interface}
\label{sec:wssetup}

@@

@@

\subsection{Setup of the A-REX with GridFTP Interface}

The A-REX can process jobs submitted through a traditional GridFTP
server instead of, or as well as, the WS interface. The GridFTP/Grid
Manager server packages can be installed alongside A-REX packages, so
it is easy to switch between the two. The configuration of the GridFTP
server is in the same legacy \emph{arc.conf} configuration file as is
used by the A-REX, and configuration instructions are given
in~\cite{gm}. In general no configuration changes are necessary to use
a Grid Manager-configured GridFTP server with the A-REX.

If it is desired to run both WS and GridFTP interfaces at the same
time, simply setup and start the A-REX as described in the previous
section, then start the GridFTP server. If only the GridFTP interface
is to be used, then the special \emph{a-rex} command should be used
(rather than \emph{arched}).

\begin{shaded}
\begin{verbatim}
$ARC_LOCATION/etc/init.d/a-rex start
\end{verbatim}
\end{shaded}

This command automatically creates a HED configuration for running the
A-REX without the WS interface, and so no manual HED configuration
needs to be done. The GridFTP server should then be started.

\begin{framed}
IMPORTANT: Do not run the A-REX and Grid Manager services at the same
time, as they will interfere with each other with unpredictable
results.
\end{framed}
